Clearance Project - 4 - PhysChem Clustering

### 2025-08-22 data-raw/ClearanceProject-4-PhysChemClustering.Rmd

@@

>>>>

```{r library-loading}
today <- format(Sys.Date(), "%d%b%Y")
devtools::load_all()
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(ctxR)
library(ggdist)
library(patchwork)
library(httk)
load_dawson2021()
```

@@

# Binding kinetics

Here what I want to do is understand the binding kinetics for available chemicals.
These don't all 
```{r Kd}
kd_halfway <- red_assay_res %>%
  select(DTXSID, Name, CAS, starts_with("Affinity.Kd.")) %>%
  distinct()
kd_halfway <- kd_halfway[apply(!is.na(kd_halfway[4:6]), MARGIN = 1, any), ]

t1.2_halfway <- kd_halfway %>%
  mutate(across(.cols = starts_with("Affinity"),
                \(x) {log(2)/(x*7E9)},
                .names = "t1.2min_{col}"),
         across(.cols = starts_with("Affinity"),
                \(x) {log(2)/(x*3.5E4)},
                .names = "t1.2max_{col}"),
         across(.cols = starts_with("Affinity"),
                \(x) {log(100)/(x*7E9)},
                .names = "t99min_{col}"),
         across(.cols = starts_with("Affinity"),
                \(x) {log(100)/(x*3.5E4)},
                .names = "t99max_{col}"))

t50_pl <- t1.2_halfway %>%
  ggplot(aes(x = forcats::fct_reorder(DTXSID, t1.2min_Affinity.Kd.Low))) +
  geom_pointinterval(aes(y = t1.2min_Affinity.Kd.Med,
                         ymin = t1.2min_Affinity.Kd.High,
                         ymax = t1.2min_Affinity.Kd.Low),
                     alpha = 0.5, shape = 20) +
  geom_pointinterval(aes(y = t1.2max_Affinity.Kd.Med,
                         ymin = t1.2max_Affinity.Kd.High, 
                         ymax = t1.2max_Affinity.Kd.Low), 
                     alpha = 0.5, shape = 20, color = "purple3") +
  scale_y_log10(labels = scales::label_math(format = log10)) + 
  labs(x = "Chemicals",y = "Binding t<sub>1/2</sub>",
       title = "t<sub>1/2</sub> = ln(2)/(K<sub>D</sub>*k<sub>on</sub>)") +
  ggtext::geom_richtext(
    x = 200, y = 1, vjust = 0,
    label = "k<sub>on,max</sub> = 7E9 M<sup>-1</sup>s<sup>-1</sup>") +
  ggtext::geom_richtext(
    x = 100, y = 1, vjust = 0,
    label = "k<sub>on,min</sub> = 3.5E4 M<sup>-1</sup>s<sup>-1</sup>", 
    color = "purple4") +
  geom_hline(yintercept = median(t1.2_halfway$t1.2min_Affinity.Kd.Med), 
             linetype = "dotted", color = "grey10") +
  geom_hline(yintercept = median(t1.2_halfway$t1.2max_Affinity.Kd.Med), 
             linetype = "dotted", color = "purple3") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title = ggtext::element_markdown(), 
        plot.title = ggtext::element_markdown(hjust = 0.5),
        panel.grid.major.x = element_blank())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Binding_halflife_toEquilibrium.png"),
       bg = "white",
       height = 5, width = 10
)

# What about 99% time
t99_pl <- t1.2_halfway %>%
  ggplot(aes(x = forcats::fct_reorder(DTXSID, t99min_Affinity.Kd.Low))) +
  geom_pointinterval(aes(y = t99min_Affinity.Kd.Med,
                         ymin = t99min_Affinity.Kd.High,
                         ymax = t99min_Affinity.Kd.Low),
                     alpha = 0.5, shape = 20) +
  geom_pointinterval(aes(y = t99max_Affinity.Kd.Med,
                         ymin = t99max_Affinity.Kd.High, 
                         ymax = t99max_Affinity.Kd.Low), 
                     alpha = 0.5, shape = 20, color = "purple3") +
  scale_y_log10(labels = scales::label_math(format = log10)) + 
  labs(x = "Chemicals",y = "Binding t<sub>0.99</sub>",
       title = "t<sub>0.99</sub> = ln(100)/(K<sub>D</sub>*k<sub>on</sub>)") +
  ggtext::geom_richtext(
    x = 200, y = 1, vjust = 0,
    label = "k<sub>on,max</sub> = 7E9 M<sup>-1</sup>s<sup>-1</sup>") +
  ggtext::geom_richtext(
    x = 100, y = 1, vjust = 0,
    label = "k<sub>on,min</sub> = 3.5E4 M<sup>-1</sup>s<sup>-1</sup>", 
    color = "purple4") +
  geom_hline(yintercept = median(t1.2_halfway$t99min_Affinity.Kd.Med), 
             linetype = "dotted", color = "grey10") +
  geom_hline(yintercept = median(t1.2_halfway$t99max_Affinity.Kd.Med), 
             linetype = "dotted", color = "purple3") +
  theme_bw() +
  theme(axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.title = ggtext::element_markdown(), 
        plot.title = ggtext::element_markdown(hjust = 0.5),
        panel.grid.major.x = element_blank())

ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Binding_99percent_toEquilibrium.png"),
       bg = "white",
       height = 5, width = 10
)

t50_pl/t99_pl
ggsave(filename = paste0(Sys.getenv("FIG_DIR"),
                         "RestrictiveClearance_Analysis/",
                         "CURRENT_", "Binding_time_toEquilibrium.png"),
       bg = "white",
       height = 10, width = 10
)
```
